---
title: "Folk"
author: "Maximiliano Garnier"
# date: "3/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(janitor)
library(RColorBrewer)
library(viridisLite)
library(ggtern)
library(plyr)
library(plotly)
library(tidyverse)
```

# Preparing the Data

Firstly, we need to create the dictionary of points.

```{r}
#Build a library of points, left to right, top to bottom...
points.Folk = vroom::vroom("data/Folk.csv", 
                           col_names = F, 
                           col_types = c(.default = "d")) %>% 
  na.omit() %>% 
  mutate(id = 1:nrow(.)) %>% 
  relocate(id)
  # select(id,X3,X1,X2)
  

colnames(points.Folk) = c("IDPoint","T","L","R")
points.Folk
```

To visualise the points, the following code can be used:

```{r}
base.Folk <- ggtern(data=points.Folk,aes(L,T,R)) +
  theme_bw() +
  geom_point(shape=21,size=5,color="blue",fill="white") +
  geom_mask() + 
  geom_text(aes(label=IDPoint),color="blue")
base.Folk
```

Assign each polygon a unique number and respective label.

```{r}
#Give each Polygon a number
polygon.labels.Folk <- data.frame(
  Label=c("Quartzarenite",
          "Subarkose",
          "Sublitharenite",
          "Arkose",
          "Lithik Arkose",
          "Feldspathic Litharenite",
          "Litharenite"),
  Label.es = c("Quarzo arenita",
          "Subarcosa",
          "Sublitarenita",
          "Arcosa",
          "Arcosa lítica",
          "Litarenita feldespática",
          "Litarenita"))
#Assign each label an index
polygon.labels.Folk$IDLabel=1:nrow(polygon.labels.Folk)
```

Create the map between the polygon numbers and the points which make up those numbers. Make sure they are in clockwise or anticlockwise order (but not mixed)

```{r}
#Create a map of polygons to points
polygons.Folk <- data.frame(
  rbind(c(1,1),c(1,2),c(1,3),c(1,4),
        c(2,2),c(2,5),c(2,6),c(2,3),
        c(3,3),c(3,6),c(3,7),c(3,4),
        c(4,5),c(4,8),c(4,9),c(4,13),
        c(5,13),c(5,9),c(5,10),c(5,6),
        c(6,6),c(6,10),c(6,11),c(6,14),
        c(7,14),c(7,11),c(7,12),c(7,7)
  )
)
#IMPORTANT FOR CORRECT ORDERING.
polygons.Folk$PointOrder <- 1:nrow(polygons.Folk)
 
#Rename the columns
colnames(polygons.Folk) = c("IDLabel","IDPoint","PointOrder")
```

Now we merge the polygons, points and polygon labels to create a master dataframe.

```{r}
#Merge the three sets together to create a master set.
df.Folk <- merge(polygons.Folk,points.Folk)
df.Folk <- merge(df.Folk,polygon.labels.Folk)
df.Folk <- df.Folk[order(df.Folk$PointOrder),]
df.Folk = df.Folk %>% 
  mutate(Label = as_factor(Label),
         Label.es = as_factor(Label.es))
```

We also create a separate data frame for the labels positioned at the centroid of each polygon.

```{r}
#Determine the Labels Data library(plyr)
Labs.Folk = ddply(df.Folk,"Label",
                 function(x){c(c(mean(x$T),mean(x$L),mean(x$R)))})
colnames(Labs.Folk) = c("Label","T","L","R")
```

Final tibble

```{r}
tb.Folk = df.Folk %>% 
  rename(Q=T,F=L,R=R) %>% 
  as_tibble() %>% 
  select(-c(1:3))
```

# ggplot

Now we can build the final plot, which employs the geom_polygon(…) and  geom_text(…) geometries and the above data-sets, we apply some transparency so the grid can be seen through the polygons, and base the drawing of the simple theme_bw(…) arrangement.

```{r}
# Folk.pal = colorRampPalette(brewer.pal(9,'Set1'))(n_distinct(df.Folk$Label))
Folk.pal = c("#777777", "#AACDC9", "#C0CF98", 
             "#2F4996", "#D3C6E2", "#EBD1C0", "#564E37")

Folk.gg <- ggtern(data=tb.Folk,aes(F,Q,R)) +
  geom_polygon(aes(fill=Label,group=Label),
               color="black",alpha=0.5) +
  # geom_text(data=Labs.Folk,aes(label=Label),size=2,color="black",
  #           # angle=c(60,60,60,0,0,0,0)
  theme_arrowdefault() +
  theme_clockwise() +
  # custom_percent("Percent") +
  scale_fill_manual(values = Folk.pal) +
  # scale_fill_brewer(palette = 'Set1') +
  labs(title="Folk",
       fill = "Sandstone",
       T="Q",
       L="F",
       R="R")
Folk.gg
```

```{r}
ggsave('figures/Folk.png',Folk.gg,width = 9,height = 6,dpi = 300)
```

# plotly

```{r}
# reusable function for creating annotation object
label <- function(txt) {
  list(
    text = txt, 
    x = 0.1, y = 1,
    ax = 0, ay = 0,
    xref = "paper", yref = "paper", 
    align = "center",
    font = list(family = "serif", size = 15, color = "white"),
    bgcolor = "#b3b3b3", bordercolor = "black", borderwidth = 2
  )
}

# reusable function for axis formatting
axis <- function(txt) {
  list(
    title = txt, tickformat = ".0%", tickfont = list(size = 10)
  )
}

Folk.ternaryAxes <- list(
  aaxis = axis("Q"), 
  baxis = axis("F"), 
  caxis = axis("R")
)
```

```{r}
Folk.ly = plot_ly(
  tb.Folk,
  a = ~Q, b = ~F, c = ~R, color = ~Label, 
  colors = Folk.pal %>% set_names(levels(tb.Folk$Label)),
  type = "scatterternary",
  fill = "toself", mode = "lines"
) %>%
  layout(
    annotations = label("Folk"), ternary = Folk.ternaryAxes,
    legend = list(title=list(text='<b> Sandstone </b>'))
  ) %>% 
  config(
    toImageButtonOptions = list(
      format = 'svg',
      filename = 'Folk',
      width = 9 * 96,
      height = 6 * 96
    )
  )
Folk.ly
```

