---
title: "Ternary diagrams"
author: "Maximiliano Garnier-Villarreal"
output: 
  bookdown::html_document2:
    theme: flatly
    number_sections: false
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    source_code: embed
    social: ['menu']
    theme: flatly # bootstrap, cerulean, flatly, united, sandstone, yeti
---

```{r setup, include=FALSE, message=FALSE}
# library(flexdashboard)
library(plotly)
library(tidyverse)

lapply(fs::dir_ls('functions/'), source)

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Diagrams

This document shows how to display and use, in **R**, different ternary diagrams that are used regularly in geosciences, these are (diagram - function):

-   QAP - `ternary_qap()`

-   QAP for gabbros - `ternary_qap_g()`

-   QAP for mafic rocks - `ternary_qap_m()`

-   QAP for ultramafic rocks - `ternary_qap_um()`

-   QFL for provenance - `ternary_qfl()`

-   QmFLt for provenance - `ternary_qmflt()`

-   AFM - `ternary_afm()`

-   Pyroclastic rocks - `ternary_pyroclastic()`

-   Folk sandstone classification - `ternary_folk()`

-   Shepard soil classification - `ternary_shepard()`

# Use

## Load functions

To load all the functions at once the following command can be used: `lapply(fs::dir_ls('functions/'), source)`.

## Static

A simple static diagram without data can be obtained by just calling the respective function. The result is a **ggplot2** object based on the **ggtern** package.

```{r}
ternary_folk()
```

## Dynamic

To obtain a dynamic diagram the `output` argument must be set to 'plotly', resulting in a **plotly** object.

```{r}
ternary_qap_m(output = 'plotly')
```

## Add your data!!

To add your own data to the diagram a new table must be created (imported), where the names of the columns must match the names of the vertices for the corresponding diagram, for this check the [Names](#tb-names) section where the respective names are given for each diagram (function). **The point totals (adding the three components) are not required to add to 100%.**

This is demostrated here by ploting two points in the QtFL diagram, first by ploting the data in a static diagram (**ggplot2**), and then ploting the same data in the dynamic diagram (**plotly**). 

In these data the first sample (point) adds to a 100 but the second one does not, showing that point made above.

```{r}
d = tibble(Qt=c(40,30),F=c(40,20),L=c(20,10))
d
```


```{r}
ternary_qfl() + 
  geom_point(data = d,
             color = 'forestgreen',
             size = 2,
             alpha = .7)
```

For the **plotly** object (diagram) the data must be mapped in the following order: `a` is the top vertix, `b` is the bottom left vertix, and `c` is th bottom right vertix. The things tha can be changed are the `data`, the `name` that is displayed in the legend, and the `marker` charcateristics; the rest is recommended not to be modified.

```{r}
ternary_qfl(output = 'plotly') %>% 
  add_trace(a = ~Qt, b = ~F, c = ~L,
            data = d,
            name = 'My data',
            marker = list(size = 8,
                          color = 'green',
                          opacity = .7),
            type = "scatterternary",
            mode = "markers",
            hovertemplate = paste0('Qt: %{a}<br>',
                                   'F: %{b}<br>',
                                   'L: %{c}'),
            inherit = F)
```

## Names {#tb-names}

These are the names that your data table must have for you to be able to plot your data in the corresponding diagram.

- `ternary_qap()`: Q, A, P

- `ternary_qap_g()`: Cpx, P, Opx

- `ternary_qap_m()`: Ol, P, Px

- `ternary_qap_um()`: Opx, Ol, Cpx

- `ternary_qfl()`: Qt, F, L

- `ternary_qmflt()`: F, Qm, Lt

- `ternary_afm()`: A, F, M

- `ternary_pyroclastic()`: Lapilli, BB, Ash

- `ternary_folk()`: F, Q, R

- `ternary_shepard()`: Sand, Clay, Silt

## Spanish (Español)

A spanish version is available for all the diagrams, to display this version the `language` argument must be set to 'es', by default is set to english ('en').

```{r}
ternary_shepard(language = 'es')
```


# Conclusion

This is very basic and quick introduction on how to use the `ternary_*` functions to plot ternary diagrams in **R**. Further customization of the diagramas can be achived by manipulating the **ggplot2** or **plotly** objects, specially when adding your own data.
