---
title: "Pyroclastic"
author: "Maximiliano Garnier"
# date: "3/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(janitor)
library(RColorBrewer)
library(viridisLite)
library(ggtern)
library(plyr)
library(plotly)
library(tidyverse)
```

# Preparing the Data

Firstly, we need to create the dictionary of points.

```{r}
#Build a library of points, left to right, top to bottom...
points.pyro = vroom::vroom("data/pyroclastic.csv", 
                           col_names = F, 
                           col_types = c(.default = "d")) %>% 
  na.omit() %>% 
  mutate(id = 1:nrow(.)) %>% 
  relocate(id)
  # select(id,X3,X1,X2)
  

colnames(points.pyro) = c("IDPoint","T","L","R")
points.pyro
```

To visualise the points, the following code can be used:

```{r}
base.pyro <- ggtern(data=points.pyro,aes(L,T,R)) +
  theme_bw() +
  geom_point(shape=21,size=5,color="blue",fill="white") +
  geom_mask() + 
  geom_text(aes(label=IDPoint),color="blue")
base.pyro
```

Assign each polygon a unique number and respective label.

```{r}
#Give each Polygon a number
polygon.labels.pyro <- data.frame(
  Label=c("Pyroclastic breccia /\nAgglomerate",
          "Lapilli-tuff breccia",
          "Lapilli-stone",
          "Lapilli tuff",
          "Tuff"),
  Label.es = c("Brecha piroclástica /\nAglomerado",
          "Toba brechosa",
          "Lapillita",
          "Toba lapillítica",
          "Toba"))
#Assign each label an index
polygon.labels.pyro$IDLabel=1:nrow(polygon.labels.pyro)
```

Create the map between the polygon numbers and the points which make up those numbers. Make sure they are in clockwise or anticlockwise order (but not mixed)

```{r}
#Create a map of polygons to points
polygons.pyro <- data.frame(
  rbind(c(1,1),c(1,2),c(1,3),c(1,1),
        c(2,2),c(2,4),c(2,5),c(2,3),c(2,2),
        c(3,4),c(3,6),c(3,7),c(3,4),
        c(4,4),c(4,7),c(4,8),c(4,5),c(4,4),
        c(5,5),c(5,8),c(5,9),c(5,5)
  )
)
#IMPORTANT FOR CORRECT ORDERING.
polygons.pyro$PointOrder <- 1:nrow(polygons.pyro)
 
#Rename the columns
colnames(polygons.pyro) = c("IDLabel","IDPoint","PointOrder")
```

Now we merge the polygons, points and polygon labels to create a master dataframe.

```{r}
#Merge the three sets together to create a master set.
df.pyro <- merge(polygons.pyro,points.pyro)
df.pyro <- merge(df.pyro,polygon.labels.pyro)
df.pyro <- df.pyro[order(df.pyro$PointOrder),]
df.pyro = df.pyro %>% 
  mutate(Label = as_factor(Label),
         Label.es = as_factor(Label.es))
```

We also create a separate data frame for the labels positioned at the centroid of each polygon.

```{r}
#Determine the Labels Data library(plyr)
Labs.pyro = ddply(df.pyro,"Label",
                 function(x){c(c(mean(x$T),mean(x$L),mean(x$R)))})
colnames(Labs.pyro) = c("Label","T","L","R")
```

Final tibble

```{r}
tb.pyro = df.pyro %>% 
  rename(BB=T,Lapilli=L,Ash=R) %>% 
  as_tibble() %>% 
  select(-c(1:3))
```

# ggplot

Now we can build the final plot, which employs the geom_polygon(…) and  geom_text(…) geometries and the above data-sets, we apply some transparency so the grid can be seen through the polygons, and base the drawing of the simple theme_bw(…) arrangement.

```{r}
# pyro.pal = colorRampPalette(brewer.pal(9,'Set1'))(n_distinct(df.pyro$Label))

pyro.gg <- ggtern(data=tb.pyro,aes(Lapilli,BB,Ash)) +
  geom_polygon(aes(fill=Label,group=Label),
               color="black",alpha=0.5) +
  # geom_text(data=Labs.pyro,aes(label=Label),size=2,color="black",
  #           # angle=c(60,60,60,0,0,0,0)
  theme_arrowdefault() +
  theme_clockwise() +
  # custom_percent("Percent") +
  # scale_fill_manual('',values = viridis(5,direction = 1,option = 'A')) +
  scale_fill_brewer(palette = 'Set1') +
  labs(title="Pyroclastic",
       fill = "",
       T="Blocks &\nBombs (> 64 mm)",
       L="Lapilli\n(2-64 mm)",
       R="Ash\n(< 2 mm)")
pyro.gg
```

```{r}
ggsave('figures/pyro.png',pyro.gg,width = 9,height = 6,dpi = 300)
```

# plotly

```{r}
# reusable function for creating annotation object
label <- function(txt) {
  list(
    text = txt, 
    x = 0.1, y = 1,
    ax = 0, ay = 0,
    xref = "paper", yref = "paper", 
    align = "center",
    font = list(family = "serif", size = 15, color = "white"),
    bgcolor = "#b3b3b3", bordercolor = "black", borderwidth = 2
  )
}

# reusable function for axis formatting
axis <- function(txt) {
  list(
    title = txt, tickformat = ".0%", tickfont = list(size = 10)
  )
}

pyro.ternaryAxes <- list(
  aaxis = axis("Blocks & Bombs (> 64 mm)"), 
  baxis = axis("Lapilli (2-64 mm)"), 
  caxis = axis("Ash (< 2 mm)")
)
```

```{r}
pyro.ly = plot_ly(
  tb.pyro,
  a = ~BB, b = ~Lapilli, c = ~Ash, color = ~Label, 
  # colors = pyro.pal %>% set_names(levels(tb.pyro$Label)),
  colors = 'Set1',
  type = "scatterternary",
  fill = "toself", 
  mode = "lines"
) %>%
  layout(
    annotations = label("Pyroclastic"), ternary = pyro.ternaryAxes
  ) %>% 
  config(
    toImageButtonOptions = list(
      format = 'svg',
      filename = 'Pyroclastic',
      width = 9 * 96,
      height = 6 * 96
    )
  )
pyro.ly
```

