---
title: "QAP-UM"
author: "Maximiliano Garnier"
# date: "3/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(janitor)
library(RColorBrewer)
library(viridisLite)
library(plotly)
library(ggtern)
library(plyr)
library(tidyverse)
```

# Preparing the Data

Firstly, we need to create the dictionary of points.

```{r}
#Build a library of points, left to right, top to bottom...
points.QAP_UM = vroom::vroom("data/QAP-UM.csv", 
                             col_names = F, 
                             col_types = c(.default = "d")) %>% 
  na.omit() %>% 
  mutate(id = 1:nrow(.)) %>% 
  relocate(id)
# select(id,X3,X1,X2)


colnames(points.QAP_UM) = c("IDPoint","T","L","R")
points.QAP_UM
```

To visualise the points, the following code can be used:

```{r}
base.QAP_UM <- ggtern(data=points.QAP_UM,aes(L,T,R)) +
  theme_bw() +
  geom_point(shape=21,size=5,color="blue",fill="white") +
  geom_mask() + 
  geom_text(aes(label=IDPoint),color="blue")
base.QAP_UM
```

Assign each polygon a unique number and respective label.

```{r}
#Give each Polygon a number
polygon.labels.QAP_UM <- data.frame(
  Label.en = c('Dunite',
               'Harzburgite',
               'Lherzolite',
               'Wehrlite',
               'Olivine orthopyroxenite',
               'Olivine websterite',
               'Olivine clinopyroxenite',
               'Orthopyroxenite',
               'Websterite',
               'Clinopyroxenite'
  ),
  Label.es = c('Dunita',
               'Harzburgita',
               'Lherzolita',
               'Wehrlita',
               'Ortopiroxenita olivínica',
               'Websterita olivínica',
               'Clinopiroxenita olivínica',
               'Ortopiroxenita',
               'Websterita',
               'Clinopiroxenita'
  ))
#Assign each label an index
polygon.labels.QAP_UM$IDLabel=1:nrow(polygon.labels.QAP_UM)
```

Create the map between the polygon numbers and the points which make up those numbers. Make sure they are in clockwise or anticlockwise order (but not mixed)

```{r}
#Create a map of polygons to points
polygons.QAP_UM <- data.frame(
  rbind(c(1,1),c(1,2),c(1,3),
        c(2,2),c(2,5),c(2,6),c(2,4),
        c(3,4),c(3,6),c(3,7),
        c(4,4),c(4,7),c(4,8),c(4,3),
        c(5,5),c(5,9),c(5,10),c(5,6),
        c(6,6),c(6,10),c(6,11),c(6,7),
        c(7,7),c(7,11),c(7,12),c(7,8),
        c(8,9),c(8,13),c(8,14),
        c(9,10),c(9,14),c(9,15),c(9,11),
        c(10,12),c(10,15),c(10,16)
  )
)
#IMPORTANT FOR CORRECT ORDERING.
polygons.QAP_UM$PointOrder <- 1:nrow(polygons.QAP_UM)

#Rename the columns
colnames(polygons.QAP_UM) = c("IDLabel","IDPoint","PointOrder")
```

Now we merge the polygons, points and polygon labels to create a master dataframe.

```{r}
#Merge the three sets together to create a master set.
df.QAP_UM <- merge(polygons.QAP_UM,points.QAP_UM)
df.QAP_UM <- merge(df.QAP_UM,polygon.labels.QAP_UM)
df.QAP_UM <- df.QAP_UM[order(df.QAP_UM$PointOrder),]
df.QAP_UM = df.QAP_UM %>% 
  mutate(Label.en = as_factor(Label.en),
         Label.es = as_factor(Label.es))
```

We also create a separate data frame for the labels positioned at the centroid of each polygon.

```{r}
#Determine the Labels Data library(plyr)
Labs.QAP_UM = ddply(df.QAP_UM,"Label.en",
                    function(x){c(c(mean(x$T),mean(x$L),mean(x$R)))})
colnames(Labs.QAP_UM) = c("Label","Q","A","P")
```

Final tibble

```{r}
tb.QAP_UM = df.QAP_UM %>% 
  rename(Ol=T,Opx=L,Cpx=R) %>% 
  as_tibble() %>% 
  select(-c(1:3))
```


# ggplot

Now we can build the final plot, which employs the geom_polygon(…) and  geom_text(…) geometries and the above data-sets, we apply some transparency so the grid can be seen through the polygons, and base the drawing of the simple theme_bw(…) arrangement.

```{r}
QAP_UM.gg <- ggtern(data=tb.QAP_UM,aes(Opx,Ol,Cpx)) +
  geom_polygon(aes(fill=Label.en,group=Label.en),
               color="black",alpha=0.5) +
  theme_arrowdefault() +
  theme_clockwise() +
  # custom_percent("Percent") +
  scale_fill_manual('',values = viridis(10,direction = -1,option = 'D')) +
  labs(title="Ultramafic",
       T="Ol",
       L="Opx",
       R="Cpx")
QAP_UM.gg
```

```{r}
ggsave('figures/QAP_UM.png',QAP_UM.gg,width = 9,height = 6,dpi = 300)
```

# plotly

```{r}
# reusable function for creating annotation object
label <- function(txt) {
  list(
    text = txt, 
    x = 0.1, y = 1,
    ax = 0, ay = 0,
    xref = "paper", yref = "paper", 
    align = "center",
    font = list(family = "serif", size = 15, color = "white"),
    bgcolor = "#b3b3b3", bordercolor = "black", borderwidth = 2
  )
}

# reusable function for axis formatting
axis <- function(txt) {
  list(
    title = txt, tickformat = ".0%", tickfont = list(size = 10)
  )
}

QAP_UM.ternaryAxes <- list(
  aaxis = axis("Ol"), 
  baxis = axis("Opx"), 
  caxis = axis("Cpx")
)
```

```{r}
QAP_UM.ly = plot_ly() %>% 
  add_trace(
    data = tb.QAP_UM,
    a = ~Ol, b = ~Opx, c = ~Cpx, 
    color = ~Label.en,
    # colors = QAP_UM.pal %>% set_names(levels(tb.QAP_UM$Label)),
    colors = viridis(10,direction = -1,option = 'D'),
    type = "scatterternary",
    fill = "toself",
    mode = "lines",
    line = list(color = "black"),
    hoverinfo = 'text',
    hoveron = 'fills'
  ) %>% 
  layout(
    annotations = label("Ultramafic"), ternary = QAP_UM.ternaryAxes
  ) %>% 
  config(
    toImageButtonOptions = list(
      format = 'svg',
      filename = 'Ultramafic',
      width = 9 * 96,
      height = 6 * 96
    )
  )
QAP_UM.ly
```


