---
title: "QAP"
author: "Maximiliano Garnier"
# date: "3/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(janitor)
library(RColorBrewer)
library(viridisLite)
library(plotly)
library(ggtern)
library(plyr)
library(tidyverse)
```

# Preparing the Data

Firstly, we need to create the dictionary of points.

```{r}
#Build a library of points, left to right, top to bottom...
points.QAP = vroom::vroom("data/QAP.csv", 
                          col_names = F, 
                          col_types = c(.default = "d")) %>% 
  na.omit() %>% 
  mutate(id = 1:nrow(.)) %>% 
  relocate(id)
  # select(id,X3,X1,X2)
  

colnames(points.QAP) = c("IDPoint","T","L","R")
points.QAP
```

To visualise the points, the following code can be used:

```{r}
base.QAP <- ggtern(data=points.QAP,aes(L,T,R)) +
  theme_bw() +
  geom_point(shape=21,size=5,color="blue",fill="white") +
  geom_mask() + 
  geom_text(aes(label=IDPoint),color="blue")
base.QAP
```

Assign each polygon a unique number and respective label.

```{r}
#Give each Polygon a number
polygon.labels.QAP <- data.frame(
  Label=c("1a",'1b','2','3a','3b',
          '4','5','6*','7*','8*',
          '9*','10*','6','7','8','9','10'),
  Plut.en = c('Quartzolite','Quartz-rich granitoid',
              'Alkali feldspar granite','Syeno-granite',
              'Monzo-granite','Granodiorite',
              'Tonalite','Quartz alkali feldspar syenite',
              'Quartz syenite','Quartz monzonite',
              'Quartz monzodiorite / Quartz monzogabbro',
              'Quartz diorite / Quartz gabbro / Quartz anorthosite',
              'Alkali feldspar syenite','Syenite','Monzonite',
              'Monzodiorite / Monzograbbro',
              'Diorite / Gabbro / Anorthosite'
  ),
  Plut.es = c('Cuarzolita','Granitoide rico en cuarzo',
              'Granito alcalifeldespático','Sieno-granito',
              'Monzo-granito','Granodiorita',
              'Tonalita','Sienita alcalifeldespática cuarzosa',
              'Sienita cuarzosa','Monzonita cuarzosa',
              'Monzodiorita cuarzosa / Monzogabro cuarzoso',
              'Diorita cuarzosa / Gabro cuarzoso / Anortosita cuarzosa',
              'Sienita alcalifeldespática','Sienita','Monzonita',
              'Monzodiorita / Monzograbro',
              'Diorita / Gabro / Anortosita'
  ),
  Volc.en = c('','',
              'Alkali feldspar rhyolite','Rhyolite',
              'Rhyodacite','Dacite',
              'Plagio-dacite','Quartz alkali feldspar trachyte',
              'Quartz trachyte','Quartz latite',
              'Quartz latite andesite / Quartz latite basalt',
              'Quartz andesite / Tholeiitic basalt',
              'Alkali feldspar trachyte','Trachyte','Latite',
              'Latite andesite / Latite basalt',
              'Andesite / Basalt'
  ),
  Volc.es = c('','',
              'Riolita alcalifeldespática','Riolita (Liparita)',
              'Riodacita','Dacita',
              'Plagidacita','Traquita cuarzo-alcalifeldespática',
              'Traquita cuarzosa','Latita cuarzosa',
              'Latiandesita cuarzosa / Latibasalto cuarzoso',
              'Andesita cuarzosa / Basalto toleítico',
              'Traquita alcalifeldespática','Traquita','Latita',
              'Latiandesita / Latibasalto',
              'Andesita / Basalto'
  ))
#Assign each label an index
polygon.labels.QAP$IDLabel=1:nrow(polygon.labels.QAP)
```

Create the map between the polygon numbers and the points which make up those numbers. Make sure they are in clockwise or anticlockwise order (but not mixed)

```{r}
#Create a map of polygons to points
polygons.QAP <- data.frame(
  rbind(c(1,1),c(1,2),c(1,3),c(1,1),
        c(2,2),c(2,4),c(2,5),c(2,3),c(2,2),
        c(3,4),c(3,10),c(3,11),c(3,6),c(3,4),
        c(4,6),c(4,11),c(4,12),c(4,7),c(4,6),
        c(5,7),c(5,12),c(5,13),c(5,8),c(5,7),
        c(6,8),c(6,13),c(6,14),c(6,9),c(6,8),
        c(7,9),c(7,14),c(7,15),c(7,5),c(7,9),
        c(8,10),c(8,16),c(8,17),c(8,11),c(8,10),
        c(9,11),c(9,17),c(9,18),c(9,12),c(9,11),
        c(10,12),c(10,18),c(10,19),c(10,13),c(10,12),
        c(11,13),c(11,19),c(11,20),c(11,14),c(11,13),
        c(12,14),c(12,20),c(12,21),c(12,15),c(12,14),
        c(13,16),c(13,22),c(13,23),c(13,17),c(13,16),
        c(14,17),c(14,23),c(14,24),c(14,18),c(14,17),
        c(15,18),c(15,24),c(15,25),c(15,19),c(15,18),
        c(16,19),c(16,25),c(16,26),c(16,20),c(16,19),
        c(17,20),c(17,26),c(17,27),c(17,21),c(17,20)
  )
)
#IMPORTANT FOR CORRECT ORDERING.
polygons.QAP$PointOrder <- 1:nrow(polygons.QAP)
 
#Rename the columns
colnames(polygons.QAP) = c("IDLabel","IDPoint","PointOrder")
```

Now we merge the polygons, points and polygon labels to create a master dataframe.

```{r}
#Merge the three sets together to create a master set.
df.QAP <- merge(polygons.QAP,points.QAP)
df.QAP <- merge(df.QAP,polygon.labels.QAP)
df.QAP <- df.QAP[order(df.QAP$PointOrder),]
df.QAP = df.QAP %>% 
  mutate(Label = as_factor(Label))
```

We also create a separate data frame for the labels positioned at the centroid of each polygon.

```{r}
#Determine the Labels Data library(plyr)
Labs.QAP = ddply(df.QAP,"Label",
                 function(x){c(c(mean(x$T),mean(x$L),mean(x$R)))})
colnames(Labs.QAP) = c("Label","Q","A","P")
```

Final tibble

```{r}
tb.QAP = df.QAP %>% 
  rename(Q=T,A=L,P=R) %>% 
  as_tibble() %>% 
  select(-c(1:3))
```


# ggplot

Now we can build the final plot, which employs the geom_polygon(…) and  geom_text(…) geometries and the above data-sets, we apply some transparency so the grid can be seen through the polygons, and base the drawing of the simple theme_bw(…) arrangement.

```{r}

QAP.gg <- ggtern(data=tb.QAP,aes(A,Q,P)) +
  geom_polygon(fill='white',aes(group=Label),
               color="black",alpha=0.05) +
  geom_text(data=Labs.QAP,aes(label=Label),size=2.5,color="black") +
  theme_bw() +
  theme_arrowdefault() +
  theme_clockwise() +
  # custom_percent("Percent") +
  labs(title="QAP",
       T="Q",
       L="A",
       R="P")
QAP.gg
```

```{r}
ggsave('figures/QAP.png',QAP.gg,width = 9,height = 6,dpi = 300)
```

# plotly

```{r}
# reusable function for creating annotation object
label <- function(txt) {
  list(
    text = txt, 
    x = 0.1, y = 1,
    ax = 0, ay = 0,
    xref = "paper", yref = "paper", 
    align = "center",
    font = list(family = "serif", size = 15, color = "white"),
    bgcolor = "#b3b3b3", bordercolor = "black", borderwidth = 2
  )
}

# reusable function for axis formatting
axis <- function(txt) {
  list(
    title = txt, tickformat = ".0%", tickfont = list(size = 10)
  )
}

QAP.ternaryAxes <- list(
  aaxis = axis("Q"), 
  baxis = axis("A"), 
  caxis = axis("P")
)
```

```{r}
QAP.ly = plot_ly() %>% 
  add_trace(
    data = tb.QAP,
    a = ~Q, b = ~A, c = ~P, 
    color = ~Label,
    # colors = QAP.pal %>% set_names(levels(tb.QAP$Label)),
    colors = 'transparent',
    type = "scatterternary",
    fill = "toself",
    mode = "lines",
    line = list(color = "black"),
    hoverinfo = 'text',
    text = ~paste('</br>', Label,
                  '</br>', Plut.es),
    hoveron = 'fills'
) %>%
  add_trace(
    data = Labs.QAP,
    a = ~Q, b = ~A, c = ~P,
    text = ~Label,
    type = "scatterternary",
    mode = "text",
    name = 'ID',
    hoverinfo = 'none',
    textfont = list(family= "Arial", size= 12, color= "Black"),
    showlegend = T
  ) %>% 
  layout(
    annotations = label("QAP"), ternary = QAP.ternaryAxes
  ) %>% 
  config(
    toImageButtonOptions = list(
      format = 'svg',
      filename = 'QAP',
      width = 9 * 96,
      height = 6 * 96
    )
  )
QAP.ly
```

# New data

```{r}
new.QAP = tibble(Q=c(23,26.9,8.5),
                 A=c(27,23.7,5.1),
                 P=c(36,39.8,59.6))
```

```{r}
QAP.gg + 
  geom_point(data = new.QAP)
```

```{r}
QAP.ly %>% 
  add_trace(a = ~Q, b = ~A, c = ~P,
            data = new.QAP,
            name = 'New',
            type = "scatterternary",
            mode = "markers",
            symbol = I('+'),
            marker = list(size=10,color='coral'),
            hovertemplate = paste0('Q: %{a}<br>',
                                   'A: %{b}<br>',
                                   'P: %{c}'))
```

