---
title: "QtFL"
author: "Maximiliano Garnier"
# date: "3/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(janitor)
library(RColorBrewer)
library(viridisLite)
library(plotly)
library(ggtern)
library(plyr)
library(tidyverse)
```

# Preparing the Data

Firstly, we need to create the dictionary of points.

```{r}
#Build a library of points, left to right, top to bottom...
points.QFL = vroom::vroom("data/QFL.csv", 
                          col_names = F, 
                          col_types = c(.default = "d")) %>% 
  na.omit() %>% 
  mutate(id = 1:nrow(.)) %>% 
  # relocate(id) 
  select(id,X3,X1,X2)
  

colnames(points.QFL) = c("IDPoint","T","L","R")
points.QFL
```

To visualise the points, the following code can be used:

```{r}
base.QFL <- ggtern(data=points.QFL,aes(L,T,R)) +
  theme_bw() +
  geom_point(shape=21,size=5,color="blue",fill="white") +
  geom_mask() + 
  geom_text(aes(label=IDPoint),color="blue")
base.QFL
```

Assign each polygon a unique number and respective label.

```{r}
#Give each Polygon a number
polygon.labels.QFL <- data.frame(
  Label=c("Craton interior",
          "Transitional Continental",
          "Basement Uplift",
          "Recycled Orogenic",
          "Dissected Arc",
          "Transitional Arc",
          "Undissected Arc"),
  Label.es = c("Cratón interior",
          "Transición continental",
          "Basamento elevado",
          "Orógeno reciclado",
          "Arco disectado",
          "Arco transicional",
          "Arco no disectado"))
#Assign each label an index
polygon.labels.QFL$IDLabel=1:nrow(polygon.labels.QFL)

polygon.labels2.QFL <- data.frame(
  Label=c("Continental Block",
          "Recycled Orogen",
          "Magmatic Arc"))
#Assign each label an index
polygon.labels2.QFL$IDLabel=1:nrow(polygon.labels2.QFL)
```

Create the map between the polygon numbers and the points which make up those numbers. Make sure they are in clockwise or anticlockwise order (but not mixed)

```{r}
#Create a map of polygons to points
polygons.QFL <- data.frame(
  rbind(c(1,1),c(1,3),c(1,4),c(1,2),
        c(2,3),c(2,5),c(2,6),c(2,4),
        c(3,5),c(3,11),c(3,12),c(3,10),c(3,6),
        c(4,2),c(4,4),c(4,6),c(4,7),c(4,8),
        c(5,6),c(5,10),c(5,7),
        c(6,10),c(6,12),c(6,13),c(6,9),c(6,8),c(6,7),
        c(7,9),c(7,13),c(7,14)
  )
)
#IMPORTANT FOR CORRECT ORDERING.
polygons.QFL$PointOrder <- 1:nrow(polygons.QFL)
 
#Rename the columns
colnames(polygons.QFL) = c("IDLabel","IDPoint","PointOrder")
```


```{r}
polygons2.QFL <- data.frame(
  rbind(c(1,1),c(1,11),c(1,12),c(1,2),
        c(2,2),c(2,6),c(2,8),
        c(3,6),c(3,12),c(3,14),c(3,8)
  )
)
#IMPORTANT FOR CORRECT ORDERING.
polygons2.QFL$PointOrder <- 1:nrow(polygons2.QFL)
 
#Rename the columns
colnames(polygons2.QFL) = c("IDLabel","IDPoint","PointOrder")
```

Now we merge the polygons, points and polygon labels to create a master dataframe.

```{r}
#Merge the three sets together to create a master set.
df.QFL <- merge(polygons.QFL,points.QFL)
df.QFL <- merge(df.QFL,polygon.labels.QFL)
df.QFL <- df.QFL[order(df.QFL$PointOrder),]
df.QFL = df.QFL %>% 
  mutate(Label = as_factor(Label),
         Label.es = as_factor(Label.es))
```

```{r}
#Merge the three sets together to create a master set.
df2.QFL <- merge(polygons2.QFL,points.QFL)
df2.QFL <- merge(df2.QFL,polygon.labels2.QFL)
df2.QFL <- df2.QFL[order(df2.QFL$PointOrder),]
df2.QFL = df2.QFL %>% 
  mutate(Label = as_factor(Label))
```

We also create a separate data frame for the labels positioned at the centroid of each polygon.

```{r}
#Determine the Labels Data library(plyr)
Labs.QFL = ddply(df.QFL,"Label",
                 function(x){c(c(mean(x$T),mean(x$L),mean(x$R)))})
colnames(Labs.QFL) = c("Label","T","L","R")
```

Final tibble

```{r}
tb.QFL = df.QFL %>% 
  rename(Qt=T,F=L,L=R) %>% 
  as_tibble() %>% 
  select(-c(1:3))
```


# ggplot

Now we can build the final plot, which employs the geom_polygon(…) and  geom_text(…) geometries and the above data-sets, we apply some transparency so the grid can be seen through the polygons, and base the drawing of the simple theme_bw(…) arrangement.

```{r}
# QFL.pal = colorRampPalette(brewer.pal(9,'Set1'))(n_distinct(df.QFL$Label))
QFL.pal = c("#989FA7", "#5A9AE1", "#2F4996", 
            "#6B6943", 
            "#BFAED2", "#DAB7A3", "#B21C3F")

QFL.gg <- ggtern(data=tb.QFL,aes(F,Qt,L)) +
  geom_polygon(aes(fill=Label,group=Label),
               color="black",alpha=0.5) +
  # geom_text(data=Labs.QFL,aes(label=Label),size=2,color="black",
  #           # angle=c(60,60,60,0,0,0,0)
  #           ) +
  # geom_polygon(aes(col=Label,group=Label),df2.QFL,
  #              fill="white",alpha=0,size=1) +
  theme_arrowdefault() +
  theme_clockwise() +
  # custom_percent("Percent") +
  scale_fill_manual(values = QFL.pal) +
  # scale_fill_brewer(palette = 'Set1') +
  labs(title="QtFL",
       fill = "Provenance",
       T="Qt",
       L="F",
       R="L")
QFL.gg
```

```{r}
ggsave('figures/QtFL.png',QFL.gg,width = 9,height = 6,dpi = 300)
```

# plotly

```{r}
# reusable function for creating annotation object
label <- function(txt) {
  list(
    text = txt, 
    x = 0.1, y = 1,
    ax = 0, ay = 0,
    xref = "paper", yref = "paper", 
    align = "center",
    font = list(family = "serif", size = 15, color = "white"),
    bgcolor = "#b3b3b3", bordercolor = "black", borderwidth = 2
  )
}

# reusable function for axis formatting
axis <- function(txt) {
  list(
    title = txt, tickformat = ".0%", tickfont = list(size = 10)
  )
}

QtFL.ternaryAxes <- list(
  aaxis = axis("Qt"), 
  baxis = axis("F"), 
  caxis = axis("L")
)
```

```{r}
QtFL.ly = plot_ly(
  tb.QFL,
  a = ~Qt, b = ~F, c = ~L, 
  color = ~Label, 
  colors = QFL.pal %>% set_names(levels(tb.QFL$Label)),
  type = "scatterternary",
  fill = "toself", 
  mode = "lines"
) %>%
  layout(
    annotations = label("QtFL"), ternary = QtFL.ternaryAxes,
    legend = list(title=list(text='<b> Provenance </b>'))
  ) %>% 
  config(
    toImageButtonOptions = list(
      format = 'svg',
      filename = 'QtFL',
      width = 9 * 96,
      height = 6 * 96
    )
  )
QtFL.ly
```

# New data

```{r}
new.QtFL = tibble(Qt=40,F=40,L=20)
```

```{r}
QFL.gg + 
  geom_point(data = new.QtFL)
```

```{r}
QtFL.ly %>% 
  add_trace(a = ~Qt, b = ~F, c = ~L,
            data = new.QtFL,
            type = "scatterternary",
            mode = "markers",
            hovertemplate = paste0('Qt: %{a}<br>',
                                   'F: %{b}<br>',
                                   'L: %{c}'),
            inherit = F)
```

