---
title: "Shepard"
author: "Maximiliano Garnier"
# date: "3/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(janitor)
library(RColorBrewer)
library(viridisLite)
library(plyr)
library(plotly)
library(ggtern)
library(tidyverse)
```

# Preparing the Data

Firstly, we need to create the dictionary of points.

```{r}
#Build a library of points, left to right, top to bottom...
points.Shepard <- data.frame(
  rbind(c( 1,1.000,0.000,0.000),
        c( 2,0.750,0.250,0.000),
        c( 3,0.750,0.125,0.125),
        c( 4,0.750,0.000,0.250),
        c( 5,0.600,0.200,0.200),
        c( 6,0.500,0.500,0.000),
        c( 7,0.500,0.000,0.500),
        c( 8,0.400,0.400,0.200),
        c( 9,0.400,0.200,0.400),
        c(10,0.250,0.750,0.000),
        c(11,0.250,0.000,0.750),
        c(12,0.200,0.600,0.200),
        c(13,0.200,0.400,0.400),
        c(14,0.200,0.200,0.600),
        c(15,0.125,0.750,0.125),
        c(16,0.125,0.125,0.750),
        c(17,0.000,1.000,0.000),
        c(18,0.000,0.750,0.250),
        c(19,0.000,0.500,0.500),
        c(20,0.000,0.250,0.750),
        c(21,0.000,0.000,1.000)
  )
)
colnames(points.Shepard) = c("IDPoint","T","L","R")
points.Shepard
```

To visualise the points, the following code can be used:

```{r}
base.Shepard <- ggtern(data=points.Shepard,aes(L,T,R)) +
  theme_bw() +
  geom_point(shape=21,size=5,color="blue",fill="white") +
  geom_text(aes(label=IDPoint),color="blue")
base.Shepard
```

Assign each polygon a unique number and respective label.

```{r}
#Give each Polygon a number
polygon.labels.Shepard <- data.frame(
  Label=c("Clay",
          "Sandy Clay",
          "Silty Clay",
          "Sand + Silt + Clay",
          "Clayey Sand",
          "Clayey Silt",
          "Sand",
          "Silty Sand",
          "Sandy Silt",
          "Silt"),
  Label.es = c("Arcilla",
               "Arcilla arenosa",
               "Arcilla limosa",
               "Arena + Limo + Arcilla",
               "Arena arcillosa",
               "Limo arcilloso",
               "Arena",
               "Arena limosa",
               "Limo arenoso",
               "Limo"))
#Assign each label an index
polygon.labels.Shepard$IDLabel=1:nrow(polygon.labels.Shepard)
```

Create the map between the polygon numbers and the points which make up those numbers. Make sure they are in clockwise or anticlockwise order (but not mixed)

```{r}
#Create a map of polygons to points
polygons.Shepard <- data.frame(
  rbind(c(1,1),c(1,2),c(1,4),
        c(2,6),c(2,2),c(2,3),c(2,5),c(2,8),
        c(3,3),c(3,4),c(3,7),c(3,9),c(3,5),
        c(4,5),c(4,14),c(4,12),
        c(5,6),c(5,8),c(5,12),c(5,15),c(5,10),
        c(6,7),c(6,11),c(6,16),c(6,14),c(6,9),
        c(7,17),c(7,10),c(7,18),
        c(8,15),c(8,12),c(8,13),c(8,19),c(8,18),
        c(9,13),c(9,14),c(9,16),c(9,20),c(9,19),
        c(10,11),c(10,21),c(10,20)
  )
)
#IMPORTANT FOR CORRECT ORDERING.
polygons.Shepard$PointOrder <- 1:nrow(polygons.Shepard)

#Rename the columns
colnames(polygons.Shepard) = c("IDLabel","IDPoint","PointOrder")
```

Now we merge the polygons, points and polygon labels to create a master dataframe.

```{r}
#Merge the three sets together to create a master set.
df.Shepard <- merge(polygons.Shepard,points.Shepard)
df.Shepard <- merge(df.Shepard,polygon.labels.Shepard)
df.Shepard <- df.Shepard[order(df.Shepard$PointOrder),]
df.Shepard = df.Shepard %>% 
  mutate(Label = as_factor(Label),
         Label.es = as_factor(Label.es))
```

We also create a separate data frame for the labels positioned at the centroid of each polygon.

```{r}
#Determine the Labels Data library(plyr)
Labs.Shepard = ddply(df.Shepard,"Label",
                     function(x){c(c(mean(x$T),mean(x$L),mean(x$R)))})
colnames(Labs.Shepard) = c("Label","T","L","R")
```

Final tibble

```{r}
tb.Shepard = df.Shepard %>% 
  rename(Clay=T,Sand=L,Silt=R) %>% 
  as_tibble() %>% 
  select(-c(1:3))
```

# ggplot

Now we can build the final plot, which employs the geom_polygon(…) and  geom_text(…) geometries and the above data-sets, we apply some transparency so the grid can be seen through the polygons, and base the drawing of the simple theme_bw(…) arrangement.

```{r}
Shepard.gg <- ggtern(data=tb.Shepard,aes(Sand,Clay,Silt)) +
  geom_polygon(aes(fill=Label,group=Label),
               color="black",alpha=0.5) +
  # geom_text(data=Labs.Shepard,aes(label=Label),
  #           size=2,color="black") +
  theme_arrowdefault() +
  theme_clockwise() +
  # custom_percent("Percent") +
  scale_fill_brewer(palette = 'Set3') + 
  labs(title="Shepard",
       fill = "Soil",
       T="Clay",
       L="Sand",
       R="Silt")
Shepard.gg
```

```{r}
ggsave('figures/Shepard.png',Shepard.gg,width = 9,height = 6,dpi = 300)
```

# plotly

```{r}
# reusable function for creating annotation object
label <- function(txt) {
  list(
    text = txt, 
    x = 0.1, y = 1,
    ax = 0, ay = 0,
    xref = "paper", yref = "paper", 
    align = "center",
    font = list(family = "serif", size = 15, color = "white"),
    bgcolor = "#b3b3b3", bordercolor = "black", borderwidth = 2
  )
}

# reusable function for axis formatting
axis <- function(txt) {
  list(
    title = txt, tickformat = ".0%", tickfont = list(size = 10)
  )
}

Shepard.ternaryAxes <- list(
  aaxis = axis("Clay"), 
  baxis = axis("Sand"), 
  caxis = axis("Silt")
)
```

```{r}
Shepard.ly = plot_ly(
  tb.Shepard,
  a = ~Clay, b = ~Sand, c = ~Silt, color = ~Label, 
  # colors = Shepard.pal %>% set_names(levels(tb.Shepard$Label)),
  colors = 'Set3',
  type = "scatterternary",
  fill = "toself", mode = "lines"
) %>%
  layout(
    annotations = label("Shepard"), ternary = Shepard.ternaryAxes,
    legend = list(title=list(text='<b> Soil </b>'))
  ) %>% 
  config(
    toImageButtonOptions = list(
      format = 'svg',
      filename = 'Shepard',
      width = 9 * 96,
      height = 6 * 96
    )
  )
Shepard.ly
```
